---
title: 'Report 3: Maroubra rocky shore'
author: "Brie Sherow"
date: "11/04/2021"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
    df_print: paged
  pdf_document:
    toc: yes
---


```{r setup, warning=FALSE, message=FALSE, results='hide', include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2) #graphing
library(tidyverse) #data manipulation
# library(kableExtra) #table layouts
library(broom) #tidy tables
library(sjPlot) #tidy tables
library(sjmisc) #tidy tables
library(sjlabelled) #tidy tables
library(car) #better qqplots with confint
library(vegan) #ordination
library(emmeans) #pairwise comparisons
library(pairwiseAdonis)

```

# Supplementary materials
The repository used to store this data can be found at [https://github.com/brieaspasia/BEES5041](https://github.com/brieaspasia/BEES5041)

```{r load-data}
#load abundance
abund <- read.csv(file = "data/multivariate/Maroubra_rock_pools.csv", header = T)

#remove X column
abund <- abund %>%
  dplyr::select(-X)
```

# Intro
In this study I will answer two questions: 1) Does gastropod abundance differ with species and height on the shore (ANOVA), and 2) Does species composition differ in rockpool vs emergent habitats, and between sampling years.  (multivariate)

# Sampling methods
  The study location is a rocky shoreline on Maroubra Beach in Sydney, Australia. 
  
  To test the hypothesis that abundance of gastropod species changes along shore gradients, students chose sites randomly along the rocky shore. An x,y grid of quadrats was created with a tape measure in a T-shape with the long end parallel to the shore. Coordinates of the samples were chosen with a random numbers table. This process was completed separately for each of the three species sampled; Austrolittorina, Nodilittorina, and Nerita. For each quadrat, students recorded the number of snails within the target species, the distance to nearest rockpool, and the zone (low, mid, or high).
  
  To test the hypothesis that species composition differs between rockpool and emergent habitats, students designated an equal number of quadrats in each habitat. For each quadrat, students recorded whether the habitat was in a rockpool, and the number of each of the target species present.


# Analysis methods
  
## Exploratory analysis
    
  Pre-analysis, data was manipulated using tidyverse package (Wickham et al 2019) in R. I used ggplot graphs as an exploratory visualisation tool, plotting abundance of each species by coastal zone. Nerita are more abundant in the mid-shore zone, while Nodilittorina are more abundant in the high-shore zone. Austrolittorina has more variation in abundance, but is still more abundant in the mid-shore.

```{r explore-data, warning=FALSE, message=FALSE}
#load zones
gastropods <- read.csv(file = "data/Linear_models/MaroubraZones.csv", header = T)
      #remove X column
      gastropods <- gastropods %>%
        dplyr::select(-X)

#visualise differences in species
      ggplot(gastropods, aes(Zone, Abundance, colour=Zone)) +
        geom_boxplot() +
        facet_wrap(~Species, scales="free_y")
      
gastro_summary <- gastropods %>%
  group_by(Species, Zone) %>% 
  summarise(mean_Abundance = mean(Abundance),
  sd_Abundance = sd(Abundance),
  SE_Abundance = sd(Abundance)/sqrt(n()))

ggplot(gastro_summary, 
       aes(Zone, mean_Abundance, fill=Zone)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean_Abundance - SE_Abundance,
   ymax = mean_Abundance + SE_Abundance), 
   width=0.2) +
  facet_wrap(~Species, scales="free_y")

```

## ANOVA

  Abundance of species were contrasted between shore gradients using an ANOVA, with abundance predicted by zone, using species as an effect. I first checked assumptions with a histogram of residuals. Finding the data skewed to the left, I found log transformation of abundance to be the best (although not perfect!) fit. In order to check normality I plotted the residuals and confidence intervals in a qq-plot using car package (Foz and Weisberg 2019). I checked the homogeneity of variance by plotting the residuals vs fitted of the ANOVA. This data is likely better suited to a GLM model, but within the parameters of this class I believe the log transformation is the best fit.
  
  In order to test the differences between species and the differences between shore heights, I ran a Tukey post-hoc test. I used emmeans package (Lenth 2021) for a more intuitive output of the comparisons within species and within zones.
  
## Multivariate

  To test the species composition between rockpools and emergent habitats, I first created a community dataframe of species abundance. I then used vegan package (Oksanen et al 2020) to standardise the abundance measures as some variables were percent cover and others were counts. I used multi-dimensional scaling to create a similarity matrix, and used stress to test conformity in the data. I then joined habitat and year data to the x-y coordinates of the ordination. I then used pairwise adonis package (Arbizu 2017) for multilevel pairwise comparisons of both habitat and year.

```{r gast-anova, warning=FALSE, message=FALSE}
#contrast abundance across zones
gast.anova <- aov(Abundance ~ Zone * Species, data = gastropods)

#check assumptions: 
    #the data is highly skewed to the left, so transformations are necessary
    hist(gast.anova$residuals)
  
    # #creating transformations from sqrt, log10, and 1/8 rt
    # gastropods <- mutate(gastropods, sqrt_abund = sqrt(Abundance+1))
    # gastropods <- mutate(gastropods, rt_abund = (Abundance+1)^(1/8))
    gastropods <- mutate(gastropods, log_abund = log10(Abundance+1))
  
    # #creating anovas to test with each transformation
    # SQgast.anova <- aov(sqrt_abund ~ Zone * Species, data = gastropods)
    # RTgast.anova <- aov(rt_abund ~ Zone * Species, data = gastropods)
    LGgast.anova <- aov(log_abund ~ Zone * Species, data = gastropods)
    
    # #checking the distribution with each transformation
    # hist(SQgast.anova$residuals)
    # hist(RTgast.anova$residuals)
    hist(LGgast.anova$residuals)
    
    # #checking the QQ plot with confidence intervals
    # # par(mfrow=c(2,2))
    # qqPlot(SQgast.anova$residuals)
    # qqPlot(RTgast.anova$residuals)
    qqPlot(LGgast.anova$residuals)

  # #looking at the residuals vs fits graph
  #   plot(SQgast.anova)
  #   plot(RTgast.anova)
    plot(LGgast.anova)

#This data is likely better suited to a GLM model, but within the parameters of this class I've chosen a log transformation as the best fit.

    #summary table
    summary(LGgast.anova)
    
#Both species and zone are significant, and there is an interaction between.

    #tukey test
    TukeyHSD(LGgast.anova)
```

```{r emmeans, warning=FALSE, message=FALSE, results='hide'}
comp1 <- emmeans(LGgast.anova, ~Species|Zone)
pairs(comp1)

comp2 <- emmeans(LGgast.anova, ~Zone|Species)
pairs(comp2)
```

#we care about things within species, differences between the three, and we want to look at the differences between shore heights.  Look at it within species and within tide heights.


```{r maroubra-rockpools, warning=FALSE, message=FALSE, results='hide'}
#load data
RockPools <- read.csv(file = "data/multivariate/Maroubra_rock_pools.csv", header = TRUE)

#create abundance matrix
RockPools_vars <- dplyr::select(RockPools, -X, -Habitat, - Year, -Replicate)

#standardise community data
com <- decostand(RockPools_vars, method= "max")

#create MDS with Bray-Curtis for species composition community data
rockpools.mds <- metaMDS(com, distance = "bray", autotransform = FALSE, trace=FALSE)

#This should ideally be less than 0.2, but this is close
rockpools.mds$stress #0.2319649

#create x-y data from matrix
data.scores <- data.frame(rockpools.mds$points)

#add habitat and year to x-y data
data.scores$habitat <- RockPools$Habitat
data.scores$year <- as.factor(RockPools$Year)
```


```{r adonis-pairwise, warning=FALSE, message=FALSE}
#Both habitat and year have significant scores, and there is a significant interaction between them
df.adonis <- adonis(com ~ data.scores$habitat*data.scores$year, permutations=999, method="bray")
df.adonis
#testing habitat dispersion
df.dis <- vegdist(com, method = "bray") # make a distance matrix
df.betadisper.hab <- betadisper(df.dis, data.scores$habitat) # test dispersion
df.betadisper.hab
permutest(df.betadisper.hab) #dispersion significant
#testing year dispersion
df.dis <- vegdist(com, method = "bray") # make a distance matrix
df.betadisper.yr <- betadisper(df.dis, data.scores$year) # test dispersion
df.betadisper.yr
permutest(df.betadisper.yr) #dispersion significant

#pairwise habitat comparisons -- differences in rockpool vs emergent
hab.pair <- pairwise.adonis(df.dis, data.scores$habitat, sim.method="bray", perm = 999, sim.function = "vegdist")
hab.pair
#pairwise year comparisons -- differences between all pairs
year.pair <- pairwise.adonis(df.dis, data.scores$year, sim.method="bray", perm = 999, sim.function = "vegdist")
year.pair
```

# Results

## ANOVA results

  ANOVA results show that abundance differs with species and height on the shore. Shore height and species type are each significant predictors in species abundance, and the interaction between the two is also significant. 
  
                Df Sum Sq Mean Sq F value   Pr(>F)    
Zone           1   1.91    1.91   10.02  0.00166 ** 
Species        2  76.75   38.38  201.62  < 2e-16 ***
Zone:Species   2  14.88    7.44   39.09 2.82e-16 ***
Residuals    414  78.80    0.19  

Tukey post-hoc tests in emmeans confirmed that the shore height is a determining factor for species abundance; both in terms of looking at comparisons within each zone, and looking at comparisons within each species.

Zone = High:
 contrast                        estimate     SE  df t.ratio p.value
 Austrolittorina - Nerita           0.925 0.0737 414 12.544  <.0001 
 Austrolittorina - Nodilittorina    0.459 0.0737 414  6.220  <.0001 
 Nerita - Nodilittorina            -0.466 0.0737 414 -6.324  <.0001 

Zone = Mid:
 contrast                        estimate     SE  df t.ratio p.value
 Austrolittorina - Nerita           0.960 0.0737 414 13.018  <.0001 
 Austrolittorina - Nodilittorina    1.274 0.0737 414 17.277  <.0001 
 Nerita - Nodilittorina             0.314 0.0737 414  4.260  0.0001 

P value adjustment: tukey method for comparing a family of 3 estimates 
Species = Austrolittorina:
 contrast   estimate     SE  df t.ratio p.value
 High - Mid   -0.418 0.0737 414 -5.672  <.0001 

Species = Nerita:
 contrast   estimate     SE  df t.ratio p.value
 High - Mid   -0.383 0.0737 414 -5.198  <.0001 

Species = Nodilittorina:
 contrast   estimate     SE  df t.ratio p.value
 High - Mid    0.397 0.0737 414  5.386  <.0001 
 
```{r ANOVA-plot, warning=FALSE, message=FALSE}

boxplot(log_abund ~ Zone * Species, data = gastropods,
        names = c("High.Aust","Mid.Aust","High.Ner","Mid.Ner","High.Nod","Mid.Nod"),
        ylab="Species abundance",
        xlab="Zone")
```

## multivariate results

  The gastropod abundance was dependent on habitat (emergent or rockpool), and year. There was an interaction between habitat and year.
  
                           Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)
habitat                    1    11.256 11.2564  39.063 0.07995  0.001
year                       4     9.543  2.3858   8.280 0.06778  0.001
habitat:year               4     2.999  0.7496   2.602 0.02130  0.001


```{r MDS-plot}
ggplot(data.scores, aes(MDS1, MDS2, color = habitat, shape=year)) + geom_point(alpha=0.8)
```
A written description of your findings (10 marks, 3-4 paragraphs)

Figures that visualise those results (with figure legend) (15 marks)

The details of statistical tests that support the conclusions you made in the text (either incorporated into the text, or in a table with an appropriate table legend). (10 marks)

### Questions: 
-- Why is the dispersion significant in betadisperser and what does that mean for the results?
-- What is a better way to plot ANOVA results in ggplot that shows something different than the data exploration?
-- Why is year and interaction with year significant in the results, but doesn't appear to be in the plot?
-- How do you write about the results without just repeating yourself a million times?

# References

Pedro Martinez Arbizu (2017). pairwiseAdonis: Pairwise Multilevel
  Comparison using Adonis. R package version 0.0.1.
  
John Fox and Sanford Weisberg (2019). An {R} Companion to Applied
  Regression, Third Edition. Thousand Oaks CA: Sage. URL:
  https://socialsciences.mcmaster.ca/jfox/Books/Companion/
  
Jari Oksanen, F. Guillaume Blanchet, Michael Friendly, Roeland Kindt,
  Pierre Legendre, Dan McGlinn, Peter R. Minchin, R. B. O'Hara, Gavin L.
  Simpson, Peter Solymos, M. Henry H. Stevens, Eduard Szoecs and Helene
  Wagner (2020). vegan: Community Ecology Package. R package version
  2.5-7. https://CRAN.R-project.org/package=vegan
  
Russell V. Lenth (2021). emmeans: Estimated Marginal Means, aka
  Least-Squares Means. R package version 1.5.5-1.
  https://CRAN.R-project.org/package=emmeans
  
Wickham et al., (2019). Welcome to the tidyverse. Journal of Open
  Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
