---
title: "LinearModels"
author: "Brie Sherow"
date: "06/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(pwr)
```
#Linear models with continuous response and predictor variables

##Q3 Interpreting linear models
```{r insect-pop-temp}
#population and temperature for four different insect groups
Bug.pop <- read.csv(file = "data/Linear_models/InsectPopTemp.csv", header = TRUE)

#testing linear model popn = a +b(Temp)
popA.lm = lm(Popna ~ Tempa, data = Bug.pop) #population described by temperature
popB.lm = lm(Popnb ~ Tempb, data = Bug.pop)
popC.lm = lm(Popnc ~ Tempc, data = Bug.pop)
popD.lm = lm(Popnd ~ Tempd, data = Bug.pop)

#summaries
summary(popA.lm)
summary(popB.lm)
summary(popC.lm)
summary(popD.lm)

#residual sum of squares
sum(summary(popA.lm)$residuals^2)
sum(summary(popB.lm)$residuals^2)
sum(summary(popC.lm)$residuals^2)
sum(summary(popD.lm)$residuals^2)

#regression sum of squares
sum((predict(popA.lm) - mean(Bug.pop$Popna))^2)
sum((predict(popB.lm) - mean(Bug.pop$Popnb))^2)
sum((predict(popC.lm) - mean(Bug.pop$Popnc))^2)
sum((predict(popD.lm) - mean(Bug.pop$Popnd))^2)

#calculate the 95% confidence interval for the slope parameter
confint(popA.lm) #confint for the slope and intercept of the regression for popA
```
##Q10
```{r swim-records}
#winning times for 100m freestyle in women's olympics
swim <- read.csv(file = "data/Linear_models/Swim.records.csv", header = TRUE)

#plot the records
ggplot(swim, aes(Year,Time)) + geom_point() + ylim(c(0,85))

#find equation for the regression line
swim.lm <- lm(Time ~ Year, data = swim)
summary(swim.lm)

#plot the line of best fit
ggplot(swim, aes(Year,Time)) + geom_point() + ylim(c(0,85)) + geom_smooth(method="lm")

#predict the time for the year 2020
predict(swim.lm, list(Year = 2020))
```
##Q15 
```{r UN-stats}
#load life expectacy vs infant mortality data
UN.pop <- read.csv(file = "data/Linear_models/Unpopstats.csv", header = TRUE)

#subset female and male populations
UN.pop.female <- filter(UN.pop, Gender == "female")
UN.pop.male <- filter(UN.pop, Gender == "male")

un.f.lm <- lm(formula = LifeExp ~ InfMort, data = UN.pop.female)
summary(un.f.lm)

un.m.lm <- lm(formula = LifeExp ~ InfMort, data = UN.pop.male)
summary(un.m.lm)
```
##Q16
```{r fire-management}
#load red gum woody debris data
debris <- read.csv(file = "data/Linear_models/coarsedebris.csv", header = TRUE)

debris.lm <- lm(Debris ~ Tree.density, data=debris)
summary(debris.lm)

```
##Q18
```{r study-times}
#load study times data
study <- read.csv(file = "data/Linear_models/studyhours.csv", header = TRUE)

study.lm <- lm(grade~hours, data=study)
summary(study.lm)

#predict the time for a perfect score
predict(study.lm, list(grade = 100))

100 = -8.554 + 11.384(hours)

100+8.554 = 108.554

108.554/11.384 = 9.535664

```

###Q3 Quiz 1 How are we meant to write this equation?  Is it not slope + intercept * Temp?  

#Testing assumptions: Examining residuals to understand patterns of variation in lm

#Q2
```{r boiling-point}
#read data
hooker <- read.csv(file = "data/Linear_models/Hooker.csv", header = TRUE)

#plot the line of best fit
ggplot(hooker, aes(Temperature,Pressure)) + geom_point() +
  geom_smooth(method="lm")

#conduct regression analysis
hooker.lm <- lm(Temperature~Pressure, data= hooker)
plot(hooker.lm)

#remove outliers does not make a difference
hooker_no.outliers <- slice(hooker, -7)
hooker.outlier.lm <- lm(Temperature~Pressure, data=hooker_no.outliers)
plot(hooker.outlier.lm)
summary(hooker.outlier.lm)

hooker <- mutate(hooker, logTemp = log10(Temperature))
hooker <- mutate(hooker, logPress = log10(Pressure))

log.hooker.lm <- lm(logTemp ~ logPress, data=hooker)
summary(log.hooker.lm)
plot(log.hooker.lm)
```

#Q4
```{r growth-vs-size}
#read data
growthrate <- read.csv(file = "data/Linear_models/GrowthRate.csv", header = TRUE)

#plot the line of best fit
ggplot(growthrate, aes(BodyWt, GrowthRate)) + geom_point() + geom_smooth(method="lm")

growthrate <- mutate(growthrate, logBodyWt = log10(BodyWt))
growthrate <- mutate(growthrate, sqrtBodyWt = sqrt(BodyWt))
growthrate <- mutate(growthrate, logGrowthRate = log10(GrowthRate))
growthrate <- mutate(growthrate, sqrtGrowthRate = sqrt(GrowthRate))

#conduct regression analysis
gr.lm <- lm(BodyWt~GrowthRate, data= growthrate)
plot(gr.lm)

log.gr.lm <- lm(logBodyWt~logGrowthRate, data= growthrate)
plot(log.gr.lm)

sqrt.gr.lm <- lm(sqrtBodyWt~sqrtGrowthRate, data= growthrate)
plot(sqrt.gr.lm)
summary(log.gr.lm)
```
#Q9
```{r brain-size}
#read data
brain <- read.csv(file = "data/Linear_models/Brain.csv", header = TRUE)

#plot the line of best fit
ggplot(brain, aes(BrainWt, BodyWt)) + geom_point() + geom_smooth(method="lm")

brain.lm <- lm(BrainWt~BodyWt, data=brain)
summary(brain.lm)
plot(brain.lm)

brain <- mutate(brain, logBrainWt = log10(BrainWt))
brain <- mutate(brain, logBodyWt = log10(BrainWt))

log.brain.lm <- lm(logBrainWt~logBodyWt, data=brain)
summary(log.brain.lm)
plot(log.brain.lm)

#plot the line of best fit
ggplot(brain, aes(logBrainWt, logBodyWt)) + geom_point() + geom_smooth(method="lm")
```
#Q12
```{r amphipod}
#read data
amphipod <- read.csv(file = "data/Linear_models/Amphipod_fecundity.csv", header = TRUE)


ggplot(amphipod, aes(Length, Eggs)) + geom_point() + geom_smooth(method="lm")

amphipod.lm <- lm(Eggs~Length, data=amphipod)
plot(amphipod.lm) 
hist(amphipod.lm$residuals)
summary(amphipod.lm)

# amphipod <- mutate(amphipod, logLength = log10(Length))
# 
# log.amphipod.lm <- lm(Eggs~logLength, data=amphipod)
# plot(log.amphipod.lm) 
# hist(log.amphipod.lm$residuals)
# summary(log.amphipod.lm)

#Length = intercept + slope(Eggs)

#Length = -25.133 + 5.639(10)
```
#Q13
```{r river-flow}
river <- read.csv(file = "data/Linear_models/riverflow.csv", header = TRUE)

ggplot(river, aes(Flow.rate, Larvae.density)) + geom_point() + geom_smooth(method="lm")

river.lm <- lm(Larvae.density~Flow.rate, data=river)
hist(river.lm$residuals)
plot(river.lm)
summary(river.lm)

```

#Q14
```{r koalas}
koalas <- read.csv(file = "data/Linear_models/koalas.csv", header = TRUE)

koalas <- mutate(koalas, years.since.fire = 2021-Year.last.fire)

ggplot(koalas, aes(years.since.fire, Koalas)) + geom_boxplot() 

ggplot(koalas, aes(years.since.fire, Koalas)) +
  geom_point() + geom_smooth(method="lm")

koalas.lm <- lm(Koalas~years.since.fire, data=koalas)
hist(koalas.lm$residuals)
plot(koalas.lm)
summary(koalas.lm)

```

###Still don't understand how to test the assumptions - how do you determine outliers or transformations?  

#Linear models 2a
```{r turtles}
turtles <- read.csv(file = "data/Linear_models/Turtles.csv", header = TRUE)

str(turtles)
turtles$Temperature <- factor(turtles$Temperature)

turtles.aov <- aov(Days ~ Temperature, data = turtles)
summary(turtles.aov)

#degrees of freedom is the number of levels of your factor -1
# residual Df = a(n-1) where a is the number of levels of factor and n is sample size (replicates per level)

turtles.lm <- lm(formula = Days ~ Temperature, data = turtles)
summary(turtles.lm)

#to determine p-value, use pf(q, df1, df2, lower.tail = FALSE) where q is the value of F, df1 is the degrees of freedom for the numerator (number of factor levels -1) and df2 is the degrees of freedom of the denominator (residual degrees of freedom)

pf(15.98, 3, 36, lower.tail = FALSE)

#a significant P value rejects the null hypothesis that hatching times are the same across all four temperatures, but does not identify which are different from each other.

#checking normality
hist(turtles.aov$residuals)

#checking homogeneity of variance
plot(turtles.aov)

#checking data
ggplot(turtles, aes(Temperature, Days)) + geom_boxplot()


qtukey(0.95, #confint
       4, #number of means
       36) #residual degrees of freedom

#critical value q for error rate of 0.05 is 3.808798

TukeyHSD(turtles.aov)
```
#Notes

##ANOVA analysis of variance
Grand mean, treatment mean of each group.  Among group variation - treatment and error. Within group variation - error. Total variation - within group variation + between group variation (Among group + within group variation).  Among group should be big and within group should be small. 

1. Calculate the grand mean.
2. Calculate each treatment mean.
3. Calculate the within group sum of squares (error).
4. Calculate df (number of measurements - number of groups)
5. Calculate mean sum of squares (mean square / degrees of freedom).
6. Calculate among group sum of squares (treatment + error).
7. Calculate mean square (divide by degrees of freedom)
8. Calculate F ratio (mean square treatment / mean square error)
9. Look up the critical value for the F ratio and decide whether to reject the null hypothesis
10. Can add error within group + error among group to get total sum of squares