---
title: "prac-01"
author: "Brie Sherow"
date: "15/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(pwr)
```

##Q1 samples and populations

```{r histogram-mean}
Population <- read.csv(file = "data/prac-01/Population.csv", header = T)

#mean of population sizes
mean(Population$Size)

#population variance
pop_var <- mean((Population$Size - mean(Population$Size))^2)
pop_var

#standard deviation
sqrt(pop_var)

#histogram of population sizes
ggplot(Population, aes(Size)) + geom_histogram()
```


```{r subset}

#subset of n random sizes
Sample5 <- sample_n(Population, 5)
mean(Sample5$Size)
pop_var5 <- mean(Sample5$Size - mean(Sample5$Size))^2
pop_var5
sqrt(pop_var5)

Sample10 <- sample_n(Population, 10)
mean(Sample10$Size)
pop_var10 <- mean(Sample10$Size - mean(Sample10$Size))^2
pop_var10
sqrt(pop_var10)

Sample20 <- sample_n(Population, 20)
mean(Sample20$Size)
pop_var20 <- mean(Sample20$Size - mean(Sample20$Size))^2
pop_var20
sqrt(pop_var20)

Sample50 <- sample_n(Population, 50)
mean(Sample50$Size)
pop_var50 <- mean(Sample50$Size - mean(Sample50$Size))^2
pop_var50
sqrt(pop_var50)
```

##Q2 confidence intervals
```{r confint}
#lower and upper confidence intervals

  #subset of 5
  mean(Sample5$Size) - (qt(0.975, df = 4)) * sd(Sample5$Size/sqrt(5))
  
  mean(Sample5$Size) + (qt(0.975, df = 4)) * sd(Sample5$Size/sqrt(5))
  
  #subset of 10
  mean(Sample10$Size) - (qt(0.975, df = 9)) * sd(Sample10$Size/sqrt(10))
  
  mean(Sample10$Size) + (qt(0.975, df = 9)) * sd(Sample10$Size/sqrt(10))
  
  #subset of 20
  mean(Sample20$Size) - (qt(0.975, df = 19)) * sd(Sample20$Size/sqrt(20))
  
  mean(Sample20$Size) + (qt(0.975, df = 19)) * sd(Sample20$Size/sqrt(20))
  
  #subset of 50
  mean(Sample50$Size) - (qt(0.975, df = 19)) * sd(Sample50$Size/sqrt(50))
  
  mean(Sample50$Size) + (qt(0.975, df = 19)) * sd(Sample50$Size/sqrt(50))
```

##Q3 one sample t-test
```{r pH}
#pH levels of a new stream 
pH <- c(7.9, 7.1, 7.0, 7.2, 7.8, 8.1, 8.0, 7.7, 7.9, 7.5)
mean(pH)

#from data to test statistic (a measure of your differences and variation)
#test statistic = mean of sample - mean of other stream / sd of sample / sqrt (sample size)
my.t <- (mean(pH) - 7.5)/(sd(pH)/sqrt(length(pH)))
my.t #0.9566892 (0.957)

#critical value of t is 2.26, our difference wasn't as big as the difference that would be associated with 0.05

#probability associated with that test statistic 
#(likelihood of your value if the null hypothesis was true)
pt(my.t, df = 9, lower.tail = FALSE)*2 #0.363715

t.test(pH, alternative = "two.sided", mu = 7.5)

#Null hypothesis is rejected, but the mean of the new stream is within the confint????  **Null hypothesis is not actually rejected?**  About 1/3 of samples within that catchment would be around my result.
```
##Q4 independent samples t-test
```{r corn}
#read data
Corn <- read.csv(file = "data/prac-01/Corn.csv", header = T)

#boxplot 
ggplot(Corn, aes(Treatment, Mass)) + geom_boxplot()

#contrast corn samples
t.test(Mass~Treatment, #mass described by treatment
       data = Corn, alternative = "two.sided", 
       var.equal = TRUE)

#df value is the degrees of freedom (n-1)
```
##Q15 one sample t-test - Why is p-value incorrect?
```{r lead}
Lead <- c(225, 341, 240, 234, 270, 334, 330, 248, 169, 336)

mean(Lead)

#confidence interval
mean(Lead) - (qt(0.975, df = 9)) * sd(Lead/sqrt(10)) 


#one sample t-test (data must be a vector), concentrations of lead should not exceed 300
t.test(Lead, mu=300, alternative = "greater")

#why is p-value incorrect???
```
##Q17 two samples t-test
```{r acid-sulphate}
#read data
AS <- read.csv(file = "data/prac-01/AcidSulphate.csv", header = T)

#boxplot 
ggplot(AS, aes(Treatment, pH)) + geom_boxplot()

#contrast corn samples
t.test(pH~Treatment, #pH described by treatment
       data = AS, alternative = "two.sided", 
       var.equal = TRUE)
```

##Q18 one sample t-test
```{r glyphosphate}
glyp <- c(874, 745, 1002, 554, 837, 901, 845, 912, 698, 712)

mean(glyp) #mean is 808

#confidence interval
mean(glyp) - (qt(0.975, df = 9)) * sd(Lead/sqrt(10)) #765.3936


#one sample t-test (data must be a vector), concentrations of lead should not exceed 300
t.test(glyp, mu = 670)

#null hypothesis is rejected
```

##Q21 two sample t-test
```{r kangaroos}
	
#read data
kanga <- read.csv(file = "data/prac-01/kangaroos.csv", header = T,
                  fileEncoding="UTF-8-BOM")

#boxplot 
ggplot(kanga, aes(salinity, pop)) + geom_boxplot()

#contrast samples
t.test(pop~salinity, #mass described by treatment
       data = kanga, alternative = "two.sided", 
       var.equal = TRUE)

```
#Q22 paired samples t-test - how to run the power analysis?
```{r nitrogen}
#testing nitrogen with a new sampling device
new <- c(101, 123, 98, 87, 110, 135, 140, 148, 112, 65)

#those same samples tested with the old device
old <- c(104, 125, 99, 86, 109, 138, 140, 147, 116, 66)

#data points are not independent so the problem is avoided by turning each pair of related measurements into a set of differences that are all independent of each other.
pair.diff <- new - old

#calculate the mean and se of the differences
diff.mean <- mean(pair.diff)

diff.se <- sd(pair.diff)/sqrt(10)

sd(pair.diff)

#calculate the test statistic
soil.t <- abs(diff.mean/diff.se)

soil.t

#calculate p-value
pt(soil.t, df = 9, lower.tail=FALSE)*2 #0.09320637

SoilN <- read.csv(file = "data/prac-01/SoilN.csv", header = T)

t.test(Nitrogen ~ Machine, data = SoilN, paired = TRUE)

#boxplot 
ggplot(SoilN, aes(Machine, Nitrogen)) + geom_boxplot()

#calculate the power analysis to find out how many sample sizes we'd need in an independent samples t-test
pwr.t.test(n = 10, d = 1.1/1.852926, sig.level = 0.05, power = NULL, type = c("two.sample"))

pwr.t.test(n = 50, d = 1.1/1.852926, sig.level = 0.05, power = NULL, type = c("two.sample"))

pwr.t.test(n = NULL, d = 1.1/1.852926, sig.level = 0.05, power = 0.8, type = c("two.sample"))
#46

trt.effect <- 1.1

sigma <- 1.852926


```
##Q25 transformations
```{r abnormal-dist-lizards}

#load data
Lizard_distribution <- read.csv(file = "data/prac-01/Lizards.csv", header = T)

#is the data normally distributed?
ggplot(Lizard_distribution, aes(Temperature)) + geom_histogram()

#is the variance similar between two samples?
ggplot(Lizard_distribution, aes(Lizard,Temperature)) + geom_boxplot()

#make sure there are no negative or zero values
Lizard_distribution <- mutate(Lizard_distribution, 
                              sqrtTemp = sqrt(Temperature),
                              logTemp = log(Temperature))

#is the data normally distributed?
ggplot(Lizard_distribution, aes(sqrtTemp)) + geom_histogram()
ggplot(Lizard_distribution, aes(logTemp)) + geom_histogram()

#is the variance similar between two samples?
ggplot(Lizard_distribution, aes(Lizard,sqrtTemp)) + geom_boxplot()
ggplot(Lizard_distribution, aes(Lizard,logTemp)) + geom_boxplot()

t.test(logTemp ~ Lizard, data = Lizard_distribution, paired = TRUE)
```
##Q28 goodness of fit
```{r finches}

black <- (70*150)/100 #pop of sample should be 105
yellow <- (1*150)/100 #pop of sample should be 1.5
red <- (29*150)/100 #pop of sample should be 43.5

pchisq(9.633, df=2, lower.tail = FALSE)

chisq.test(x = c(96, 49, 5), p = c(0.7, 0.29, 0.01)) # ignore the warning sign

#testing whether 30 males and 19 females is significant
chisq.test(x = c(35, 14), p = c(0.5,0.5))

pchisq(2.4694, df=1, lower.tail = FALSE)
```

##Q32 contingency tables
```{r drains}
Drain <- matrix(c(12,12,14,10,21,3), nrow=2)

Drain

chisq.test(Drain, correct=F)
```
##Q36 paired t-test 
```{r crab}
#load crab
crab <- read.csv(file="data/prac-01/crab.csv", header=T)

t.test(Duration ~ Side, data = crab, paired = TRUE)

#boxplot 
ggplot(crab, aes(Side, Duration)) + geom_boxplot()

#The p-value is significant at 0.02, so we can reject the null hypothesis (fights between males with claws on the same side last longer than those between males with claws on opposite sides), and determine that fights between males with claws on the opposite sides last longer.  BUT this is not what the data shows??
```
##Q37 goodness of fit - how to calculate the test statistic?
```{r soccer}
#chi-squared = (observed - expected)^2 / expected

observed <- c(153,20,138)
expected <- c(100,89,122)
diff <- observed-expected

chi = (diff)^2 / expected

#expected
100*100/311  # 32.15 left kicks
89*100/311  # 28.62 centre kicks
122*100/311  # 39.23 right kicks

pchisq(9.633, df=2, lower.tail = FALSE)

pchisq(83.67, df=2 ,lower.tail = FALSE)

chisq.test(x = c(153, 20, 138), p = c(.3215, .2862, .3923)) 


```
##Q39 contingency table
```{r gecko}
gecko <- matrix(c(8,7,18,24,19,16,15,15,18,22,12,6), nrow=4)

rownames(gecko) <- c("House gecko", "Barking gecko", "Snow skink", "Water skink")
colnames(gecko) <- c("Sand", "Leaf litter", "Rock")

chisq.test(gecko, correct = F)

#checking which combination of lizard species and habitat type show the greatest deviation from expected under null (no association between species and habitat type)
chisq.test(gecko)$residuals 
```

##Q41 goodness of fit
```{r windfarms}

windfarm <- data.frame(c(12,33,59))

colnames(windfarm) <- c("Number")
rownames(windfarm) <- c("Small", "Medium", "Large")

chisq.test(x = c(12,22,59), p = c(.33, .33, .34))

chisq.test(x = c(12,22,59), p = c(.33, .33, .34))$expected


```
##Q43 contingency table
```{r opals}
opals <- matrix(c(35, 47, 18, 5, 7, 8), nrow=2)

colnames(opals) <- c("Black", "Dark", "Light")
rownames(opals) <- c("NSW", "Queensland")

a = chisq.test(opals)

chisq.test(opals)$residuals

pchisq(opals, df=1,lower.tail = FALSE)
```
##Q45 paired t-test
```{r cows}
#load cows
cows <- read.csv(file="data/prac-01/cows.csv", header=T)

#boxplot 
ggplot(cows, aes(copper, metals)) + geom_boxplot()

t.test(copper ~ metals, data = cows, paired = TRUE)

```
##Q47 goodness of fit
```{r toxin}

toxin <- (c(12,18,6,4))


chisq.test(x = c(12,18,6,4), p = c(.25, .25, .25, .25))

```
#Q49 two sample t-test
```{r corals}
#load corals
corals <- read.csv(file="data/prac-01/corals.csv", header=T)

#visualise the differences in growth between the two treatments 
  #create a column for growth
  corals <- corals %>%
    mutate(growth=size_3months-size_initial)
  #boxplot
  ggplot(corals, aes(pH_Treatment, growth)) + geom_boxplot()

#contrast samples
t.test(growth~pH_Treatment, #growth described by treatment
       data = corals, alternative = "two.sided", 
       var.equal = TRUE)

#create a vector of low values, one sample t-tes

t.test(low, mu=0)
```

##Q50 two sample t-test
```{r weeds}

before <- c(45, 37, 23, 12, 38, 24, 18, 5, 27, 8)
after <- c(35, 38, 20, 11, 25, 12, 22, 8, 28, 12)

#load weeds
weeds <- read.csv(file="data/prac-01/weeds.csv", header=T)

weeds_long <- gather(weeds, time, pct_weeds, Before:After, factor_key=TRUE)


  #boxplot
  ggplot(weeds_long, aes(time, pct_weeds)) + geom_boxplot()

t.test(x = before, y = after, paired = TRUE)

t.test(pct_weeds ~ time, data = weeds_long, paired = TRUE)
```
##Q51 goodness of fit
```{r koalas}

koalas <- read.csv(file="data/prac-01/koalas.csv", header=T)

koalas_chi <- chisq.test(x = c(8,14,2,7,24,1), p = c(.05,.25,.24,.31,.10,.05)) # ignore the warning sign

koalas_chi$residuals
```
#Notes

### Finding the population variance
  pop_var <- mean((Population$Size - mean(Population$Size))^2)
        
### Finding the standard deviation
  standard deviation <- sqrt(pop_var)
  standard_error <- sd(Population$Size) / sqrt(length(200))

### Test statistic
  sample mean - population mean / SE
  standard_error <- sd(x) / sqrt(length(x))
  my.t <- (mean(sample) - pop_mean)/(sd(sample)/sqrt(length(sample)))
  
### Assumptions of parametric tests
    independence of replicate measures (an issue of experimental design)
    errors that are normally distributed
    when contrasting groups, homogeneity of variation
  
##One sample t-test (pH, lead, glyphosphate)
  comparing one sample to a null hypothesis (everything is the same)
  t.test(sample, alternative = "two.sided", mu = null_mean) 
  
##Independent samples t-test (corn, acid sulphate, kangaroos, corals)
  comparing two separate sets of samples with two different treatments
  t.test(Measurement~Treatment, data = sample, alternative = "two.sided", var.equal = TRUE)

##Paired samples t-test (nitrogen, crab, cows, weeds)
  comparing one set of samples with two separate treatments
  pair.diff <- new - old
  diff.mean <- mean(pair.diff)
  diff.se <- sd(pair.diff)/sqrt(sample_size)
  diff.mean <- mean(pair.diff)
  t-statistic <- abs(diff.mean/diff.se)
  pt(t-statistic, df, lower.tail=FALSE)*2
  OR
  t.test(Measurement ~ Treatment, data = sample, paired = TRUE)
  
##Transformations (lizards)
  transform to log or sqrt scale to ensure normal distribution
  data <- mutate(data, log = log(Measurement))
  
##Goodness of Fit test (finches, soccer, windfarms, toxins, koalas)
  using one categorical (rather than continuous) variable with chi-squared tests
  chi-squared = (observed - expected)^2 / expected
  pchisq(chi-squared, degrees_freedom, lower.tail = FALSE)
  OR
  chisq.test(x = c(observed list), p = c(probability of expected list))
  
##Contingency tables (drains, gecko)
  using multiple categorical variables with chi-squared tests
  expected values = column total * row total / grand total *** REVIEW THIS ***
  chisq.test(sample, correct=F)
        
#Questions
### What is the difference between SE, SD, and population variance? Why is SE and SD the same in this example
### Q22 How many soil samples would be required if you wanted to conduct an independent samples t -test to test whether the two machines differed?  Is this to do with a power analysis (if so, how do you do one) or is this just suggesting to measure each sample once instead of twice?
### Q29 I don't understand this equation.  What is the sigma?
## Q34 I can't get the correct t-statistic or p-value using a calculator, I don't understand this
## Q37 I don't know how you can find one chi-squared from three values from the equation?
## Q41 test statistic is wrong, is it because of the probabilities?  I have them equal, cause that's the null?
## Q44 Null hypothesis was rejected, so the two states have different colours of opals. Q44 is not what we're testing, right?
# Q45 How is the test statistic wrong?  And I don't understand how to improve the experiment...
# Q49 Don't understand what test would test hypothesis that corals grow at a lower pH.
# Q50 How is the test statistic wrong?
# Q53 How could the test be improved?

```{r}

```




